Commutativity of operation-based CRDTs \\ CIS 673 Project Report

### 2020-04-18 report/draft.tex
Finalize report draft

\TODO{(Maybe -- If we have time) Figure out what we have to do with
  effect preconditions. Maybe we need to encode them with ite to make
  sure that if they fail, it means that an update is not enabled when
  it should be?}

@@

In order to prove that the LWW Register is Strongly Eventually
Consistent, we need to prove that all different concurrent effects
(indicated as ``downstream'' in \cref{fig:lww-def}) commute. Since
only the value $x$ of the register can be returned using queries, we
consider two states to be equivalent if they have the same value $x$.

We started by encoding the state as is (a value $x$ of any type $X$
and a timestamp $t$ of type $Int$) and by encoding the $assign$ effect
by only requiring as a precondition that $t' \neq t$, in order to
enforce timestamp uniqueness. Servois quickly retuned the following
counterexample:

\[
t_1' = t \vee t_2' = t
\]

\TODO{Gautam: Do you understand why exactly does this fail? It is in
  line 36,37 of lwwreg.yml} Servois found that if one of the effect
timestamps $t_1', t_2'$ is equal to the initial timestamp $t$, then
executing them in different orders will lead to one of them not
satisfying the precondition once, but satisfying it in the other
run. However, this behaviour is disallowed by the LWW specification
since all new timestamps $t'$ are unique.

In order to deal with that, we extended the state to keep a set of all
seen timestamps, and we modified the precondition to require that
\TODO{Finish this.}

\TODO{Describe how we implemented it, and why we need to have a set of seen Timestamps to enforce uniqueness.}

>>>>

%% \TODO{(Maybe -- If we have time) Figure out what we have to do with
%%   effect preconditions. Maybe we need to encode them with ite to make
%%   sure that if they fail, it means that an update is not enabled when
%%   it should be?}

@@

In order to prove that the LWW Register is Strongly Eventually
Consistent, we need to prove that all different concurrent effects
(indicated as ``downstream'' in \cref{fig:lww-def}) commute. We
consider the Since only the value $x$ of the register can be returned
using queries, we consider two states to be equivalent if their values
$x$ and timestamps $t$ are equal.

We started by encoding the state as a value $x$ of any type $X$ and a
timestamp $t$ of type $Int$. In order to encode the \texttt{effect} of
the $assign$ method we need to ensure that the timestamp $t'$ is
unique. In order to achieve that, we need to extend the state with an
additional component $SeenTimestamps$ that contains the set of all
timestamps that have been encountered so far. This auxiliary component
can be used to encode uniqueness as the \texttt{effect}-method
precondition:

\[
t' \not \in SeenTimestamps \wedge t \in SeenTimestamps
\]

\noindent
This ensures that all new timestamps cannot be in the set of seen
timestamps and that the current timestamp is in the set (which is an
invariant). We then specified the effect method postconditions as seen
in its specification. Servois successfully verified that the effect
method commutes with itself.
